Backend developer guide for MS data

### 2023-09-25 dev/demo_ozonation_project.Rmd
Support for ion mobility data with get_spectra(), make_ms_targets() and others
Also added Rmarkdown files for ozonation demo and backend developer guide

# Add logo file
path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
logo <- file.path(path_extdata, "logo_StreamFind.png") #207 x 216 px
```

```{r logo, echo=FALSE}
if (!is.na(logo)) {
  htmltools::img(
  src = knitr::image_uri(logo),
  alt = "logo",
  style =
    paste0("
      margin-left:inherit;
      position:absolute;
      top:100px;
      left:610px;
      width:", 250, "px;
      height:", 250*(1633/3070), "px;"
    )
  ) 
} # height:", 100*(216/207), "px;"
```

@@

The workflow can be run by chaining all the modules as demonstrated below or by running the method `run_workflow()`.

```{r do-workflow, results='markup', warning=FALSE}

ms$run_workflow()

ms

#ms$find_features()$annotate_features()$group_features()$filter_features()
```

>>>>

# Add logo file
path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
logo <- file.path(path_extdata, "logo_StreamFind.png")
```

```{r logo, echo=FALSE}
if (!is.na(logo)) {
  htmltools::img(
  src = knitr::image_uri(logo),
  alt = "logo",
  style =
    paste0("
      margin-left:inherit;
      position:absolute;
      top:150px;
      left:600px;
      width:", 250, "px;
      height:", 250*(1633/3070), "px;"
    )
  )
}
```

@@

The modules assidadded *ProcessingSettings* can be processed by `run_workflow()` as demonstrated below or by chaining all the modules. With `run_workflow()`, the processing modules are applied with the same order they were added. By chaining the order can be modified as the modules are individually called, as shown below.

```{r do-workflow, results='markup', warning=FALSE}
# Run all ProcessingSettings added to the MassSpecData
ms$run_workflow()
```

```{r do-workflow-chain, eval=FALSE}
# Alternatively, the modules can be chained
ms$find_features()$annotate_features()$group_features()$filter_features()
```

### 2023-09-25 dev/developer_guide_backend.Rmd
Support for ion mobility data with get_spectra(), make_ms_targets() and others
Also added Rmarkdown files for ozonation demo and backend developer guide

---
title: 'StreamFind: Wastewater Ozonation'
subtitle: Mass spectrometry project example
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
  affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

@@

# Create a *MassSpecData*

An [R6](https://r6.r-lib.org/) [MassSpecData](https://odea-project.github.io/StreamFind/reference/MassSpecData.html) class object is created using the `MassSpecData$new()`, as shown below for a set of *mzML* files. The original vendor files were converted to *mzML* format using the [convert_ms_files](https://odea-project.github.io/StreamFind/reference/convert_ms_files.html) function, which uses the *msConvert* command line from [ProteoWizard](https://proteowizard.sourceforge.io/download.html). The files are from influent and effluent of a wastewater ozonation step measured in positive and negative ionization mode. Blank measurements for each ionization were also included for background subtraction.

```{r files, results='markup'}
basename(files)
```

```{r create-ms, message=FALSE, warning=FALSE}
# Create a MassSpecData from mzML files
ms <- MassSpecData$new(files = files)
```

```{r show-ms, results='markup'}
# Print in console a summary of the MassSpecData
ms
```

## Add *ProjectHeaders*

Project headers (e.g., name, author and description) can be added to the MassSpecData as shown below. The headers are converted to an S3 [ProjectHeaders](https://odea-project.github.io/StreamFind/reference/ProjectHeaders.html) class object.

```{r add-headers, results='markup'}
# Add headers to the MassSpecData
ms$add_headers(
  name = "Wastewater Ozonation",
  author = "Ricardo Cunha",
  description = "StreamFind demonstration project"
)

# Get the headers as ProjectHeaders
ms$get_headers()

# Get the MassSpecData name
ms$get_headers(value = "name")
```

## Associate blanks

The analysis replicate names and the associated blank replicate name can be ammended in the MassSpecData as shown below.

```{r add-replicates, results='markup'}

# Character vector with analysis replicate names
rpls <- c(
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("in_neg", 3),
  rep("in_pos", 3),
  rep("out_neg", 3),
  rep("out_pos", 3)
)

# Character vector with associated blank replicate names
# Note that the order should match the respective replicate
blks <- c(
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("blank_neg", 3),
  rep("blank_pos", 3)
)

# Chaining add replicates and blanks
ms$add_replicate_names(rpls)$add_blank_names(blks)

# Get the blank replicate names
ms$get_blank_names()
```

# Data processing

Data processing is performed by modules according to [ProcessingSettings](https://odea-project.github.io/StreamFind/reference/ProcessingSettings.html) as shown in the following sub-chapters.

## Obtain *ProcessingSettings*

The S3 *ProcessingSettings* class objects for processing modules are obtained via the respective [Settings](https://odea-project.github.io/StreamFind/reference/index.html#processing-settings-for-ms-data) functions. Below we obtain the *ProcessingSettings* for the module `find_features()` using the algorithm *xcms3_centwave*. The parameters for each processing module can be changed via the function arguments and the information about the meaning of each parameter can be found in the [StreamFind reference documentation](https://odea-project.github.io/StreamFind/reference/index.html).

```{r ffs, results='markup'}
# Get ProcessingSettings for finding features using the xcms3 centwave algorithm
ffs <- Settings_find_features_xcms3_centwave()

# Print in console the details of the ProcessingSettings
ffs
```

```{r other-settings, results='markup'}
fls <- Settings_filter_features_StreamFind(
  minIntensity = 5000,
  minSnRatio = 20,
  maxGroupSd = 30,
  blank = 5,
  minGroupAbundance = 3,
  excludeIsotopes = TRUE
)

fls

afs <- Settings_annotate_features_StreamFind()

afs

gfs <- Settings_group_features_xcms3_peakdensity()

gfs
```

## Add *ProcessingSettings*

When all settings are available, they can be added or used directly to run the workflow. In the example below, the different *ProcessingSettings* are first added to the *MassSpecData* in the desired order.

```{r add-settings, results='markup'}
# Add settings for finding, annotating, grouping and finally filter features
ms$add_settings(settings = list(ffs, afs, gfs, fls))

# Getting the workflow settings
ms$get_settings_names()
```

## Apply the workflow

The modules assidadded *ProcessingSettings* can be processed by `run_workflow()` as demonstrated below or by chaining all the modules. With `run_workflow()`, the processing modules are applied with the same order they were added. By chaining the order can be modified as the modules are individually called, as shown below.

```{r do-workflow, results='markup', warning=FALSE}
# Run all ProcessingSettings added to the MassSpecData
ms$run_workflow()
```

```{r do-workflow-chain, eval=FALSE}
# Alternatively, the modules can be chained
ms$find_features()$annotate_features()$group_features()$filter_features()
```

# Instructions

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
This section gives quick instructions to display text, code, tables and figures.

## Text

Text in R markdown has several inline codes for costumization.
The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.

## Code

Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
Code chunks should always be unically named.

```{r exampleCode}

#Example of code chunk

list(a = 1:6, b = 1:10)

```

## Tables

The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exampleTable). Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
Table caption is added with the argument `caption` of the `kable` function.

```{r exampleTable}

df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)

knitr::kable(df, caption = "Example of table caption.") %>%
kable_styling(
  font_size = 12,
  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
  fixed_thead = TRUE
) %>%
  scroll_box(width = "100%", height = "350px")

```

## Figures

Figures/plots can also by embedded using diverse packages.
The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.

### Example `ggplot2`

```{r examplePlot, fig.cap="Plot caption example"}

ggplot(pressure, aes(temperature, pressure)) +
geom_point() +
theme_bw()

```

### Example `plotly`

```{r examplePlotly, fig.cap="Plot caption example"}

plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")

```

@@

# Signature

The report was produced on `r Sys.time()`. The data and report are stored in ``r path_wd`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.

>>>>

---
title: 'Backend developer guide for MS data'
subtitle: |
  | Implementation of backend processing modules and algorithms for mass spectrometry
  | <br>

author:
  name: Ricardo Cunha
  email: cunha@iuta.de
  affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

@@

# StreamFind

The backend of StreamFind is an R package and can be used for data management, processing, visualization and reporting. This guide focuses on mass spectrometry (MS) data and aims to instruct developers to implement new processing modules and additional processing algorithms for new or existing processing modules in StreamFind.

## Setup

The R package is in the GitHub [StreamFind repository](https://github.com/odea-project/StreamFind) of the [ODEA Project](https://github.com/odea-project). For development, the recommendation is to download the repository locally using [git](https://git-scm.com/) version control. The [GitHub desktop](https://desktop.github.com/) tool can be used for more easily install and configure git with your GitHub account, which is recommended for authoring contributions. Since it is an R package, the [RStudio](https://posit.co/products/open-source/rstudio/) IDE is recommended for development. Yet, others (e.g., [VS Code](https://code.visualstudio.com/download)) will also work. When using RStudio, the repository can be downloaded via new project, selecting version control, then git and finally adding the GitHub *url* of the StreamFind: https://github.com/odea-project/StreamFind. This should create a local image of the GitHub StreamFind repository directly with git tracking if git or GitHub desktop were properly installed and configured.

## Development

Considering that the local image of the StreamFind repository is installed with git version tracking, the first step for development is to create a dedicated branch for implementation of new processing modules and/or algorithms. The master branch should not be changed directly but modified by pull requests from the dedicated development branch, giving the opportunity for code revision.

## Structure

## Semantics

# R6 and S3

# *ProcessingSettings*

## Constructor

## S3 methods

# New module

# New *ProcessingSettings*

@@
